---
title: "Lab2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Task

Read the data
```{r}
data <- read.csv("./mlbootcamp5_train (1).csv", sep = ";")
head(data, 10)
```

Transform the features: 
- create "age in years" (`age_years`)dividing age by 365.25 and taking floor ($\lfloor{x}\rfloor$ is the largest integer that is less than or equal to $x$)
- remove "age" feature
- transfrom `cholesterol` and `gluc` to factor

```{r}
data$age_years <- floor(data$age/365.25)
data <- subset(data, select=-c(age, id))
data$cholesterol <- as.factor(data$cholesterol)
data$gluc <- as.factor(data$gluc)
data$cardio <- as.factor(data$cardio)
head(data, 10)
print(is.factor(data$cholesterol))
print(is.factor(data$gluc))
```

Split data into train and test parts in the proportion of 7/3.
The target feature is `cardio`

```{r}
library(caret)
set.seed(2344)
trainIndex <- createDataPartition(data$cardio, p = 0.7, list = FALSE, times = 1)
train <- data[trainIndex, ]
test  <- data[-trainIndex, ]
head(train, 5)
head(test, 5)
```

Train a decision tree on the dataset `training` with **max depth equal to 3**. For setting this parameter use function `rpart::rpart.control`. Use default cross calidation parameters.

```{r}
library(rpart)
model <- train(cardio ~ ., data = train, method = "rpart", control = rpart.control(maxdepth = 3))
```

Plot final tree with `rattle` library

```{r}
library(rattle)
fancyRpartPlot(model$finalModel, main="cardio")
```

What is the accuracy of prediction on test data?

```{r}
prediction <- predict(model, test)
accuracy <- sum(prediction == test$cardio) / nrow(test)
print(accuracy)
```
Now set cross validation parameters: `method = "cv"`, `number = 5`. Tree parameters set to default.

```{r}
fitControl <- trainControl(method = "cv", number = 5)
model2 <- train(cardio ~ ., data = train, method = "rpart", trControl = fitControl)
```

Plot final tree with `rattle` library

```{r}
library(rattle)
fancyRpartPlot(model2$finalModel, main="cardio")
```

What is the accuracy of prediction on test data?

```{r}
prediction2 <- predict(model2, test)
newAccuracy <- sum(prediction2 == test$cardio) / nrow(test)
print(newAccuracy)
```

Does the accuracy became better or not?
```
Точність залишилися без змін.
```