---
title: "Lab 3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Task

Read the data. Pay attention that data has NA values. It's better to convert target variable to factor.
```{r}
df<-read.csv("./donner_party.csv")
df<-na.omit(df)
df$Survived<-as.factor(df$Survived)
```

For prediction we will use only variables `Sex` and `Age`. For predicting binary outcome we will use **Generalized Linear Model** (`method = "glm"`) and **caret package**. 

```{r}
library(caret)
fit<-train(
  Survived~Sex+Age,
  data=df,
  method='glm',
  family=binomial
)
pred<-predict(fit,newdata=df)
```

What is your in-sample accuracy? Build a confusion matrix

```{r}
confusionMatrix(pred, df$Survived)
```

Use a cross-validation with `method = "repeatedcv"`, `number = 5`, `repeats = 10`. Does cross-validation improve the accuracy?

```{r}
fit2<-train(
  Survived~Sex+Age,
  data=df,
  method='glm',
  family=binomial,
  trControl=trainControl(method='repeatedcv',number=5,repeats=10)
)

pred2<-predict(
  fit2,
  newdata=df
)
confusionMatrix(pred2,df$Survived)
```

## Точність не покращилась.

